---
title: "chromeleon_untargeted_analysis"
date: 'Last updated: `r Sys.Date()`'
author: "Ruth Williamson"
format: 
  html:
    toc: true
    toc_depth: 3
    toc_float: true
    embed-resources: true
---

# Load in Libraries
```{r}
library(tidyverse)
library(cowplot)
library(readxl)
theme_set(theme_cowplot())
library(gt)
library(openxlsx)
library(lattice)
library(janitor)
library(snakecase)
library(chemhelper)
library(fuzzyjoin)
```

# Reading in test file 1
```{r}
f4hr_p14 <- read.xlsx("forest4hr_p14_extractedion.xlsx", "Sheet2", na.strings = "n.a.") |>
  slice(-n()) |> # removes row with totals of columns
  na.omit() # removes rows with NA that that were not integrating and needed to be removed
```

> General information gathered from this import:
+ There are 87 unique compounds represented

# Reading in test file 2
```{r}
f4hr_p14_tic <- read.xlsx("forest4hr_p14_tic.xlsx", "Sheet2", na.strings = "n.a.") |>
   slice(-n()) # removes row with totals of columns
```

> General information gathered from this import:
+ There are 87 unique compounds represented

# Reading in Alkane Info
```{r}
nrt <- read.xlsx("Alkane_RT_fto.xlsx")
```


# Retention Index Calculations

Retention indexes are calculated by using the retention time of alkanes that proceed directly before and after the compound.The equation is as follows

> Ix = 100n + 100(tx-tn) / (tn+1 − tn)

The package I would like to use for this is: chemhelper by Eric R. Scott

To cite package ‘chemhelper’ in publications use:

Scott E (2023). chemhelper: Helper Functions For Dealing With GCMS and LCMS data from IonAnalytics. R package version 0.0.0.9000, https://github.com/aariq/chemhelper.

Testing one retention index and comparing it to hand calculation
```{r}
calc_RI(8.551333, nrt$rt, nrt$carbon_num)

#my hand calculation resulted in 884.6
```

```{r}
f4hr_p14_tic_ri <- f4hr_p14_tic |>
  mutate(ri = calc_RI(RetentionTime, nrt$rt, nrt$carbon_num))
```

> 20 of the compounds fell outside the alkane range and were mathematically extrapolated using the chemhelper function.

```{r}
f4hr_p14_ri <- f4hr_p14 |>
  mutate(ri = calc_RI(RetentionTime, nrt$rt, nrt$carbon_num))
```

> 16 of the compounds fell outside the alkane range and were mathematically extrapolated using the chemhelper function.

