---
title: "chromeleon_untargeted_analysis"
date: 'Last updated: `r Sys.Date()`'
author: "Ruth Williamson"
format: 
  html:
    toc: true
    toc_depth: 3
    toc_float: true
    embed-resources: true
---

# Load in Libraries
```{r}
library(tidyverse)
library(cowplot)
library(readxl)
theme_set(theme_cowplot())
library(gt)
library(openxlsx)
library(lattice)
library(janitor)
library(snakecase)
library(chemhelper)
library(fuzzyjoin)
library(webchem)
```

# Reading in test file 1
```{r}
f4hr_p14 <- read.xlsx("forest4hr_p14_extractedion.xlsx", "Sheet2", na.strings = "n.a.") |>
  slice(-n()) |> # removes row with totals of columns
  na.omit() # removes rows with NA that that were not integrating and needed to be removed
```

> General information gathered from this import:
+ There are 87 unique compounds represented

# Reading in test file 2
```{r}
f4hr_p14_tic <- read.xlsx("forest4hr_p14_tic.xlsx", "Sheet2", na.strings = "n.a.") |>
   slice(-n()) # removes row with totals of columns
```

> General information gathered from this import:
+ There are 87 unique compounds represented

# Reading in Alkane Info
```{r}
nrt <- read.xlsx("Alkane_RT_fto.xlsx")
```


# Retention Index Calculations

Retention indexes are calculated by using the retention time of alkanes that proceed directly before and after the compound.The equation is as follows

> Ix = 100n + 100(tx-tn) / (tn+1 − tn)

The package I would like to use for this is: chemhelper by Eric R. Scott

To cite package ‘chemhelper’ in publications use:

Scott E (2023). chemhelper: Helper Functions For Dealing With GCMS and LCMS data from IonAnalytics. R package version 0.0.0.9000, https://github.com/aariq/chemhelper.

Testing one retention index and comparing it to hand calculation
```{r}
calc_RI(8.551333, nrt$rt, nrt$carbon_num)

#my hand calculation resulted in 884.6
```

```{r}
f4hr_p14_tic_ri <- f4hr_p14_tic |>
  mutate(ri = calc_RI(RetentionTime, nrt$rt, nrt$carbon_num))
```

> 20 of the compounds fell outside the alkane range and were mathematically extrapolated using the chemhelper function.

```{r}
f4hr_p14_ri <- f4hr_p14 |>
  mutate(ri = calc_RI(RetentionTime, nrt$rt, nrt$carbon_num))
```

> 16 of the compounds fell outside the alkane range and were mathematically extrapolated using the chemhelper function.

#Testing webchem
```{r}
CASs <- c("83-34-1", "119-36-8", "123-35-3", "19700-21-1")

RIs <-
  nist_ri(
    query = CASs,
    from = "cas",
    type = "linear",
    polarity = "non-polar",
    temp_prog = "ramp"
  )
head(RIs)

library(stringr)
RIs_filtered <-
  RIs %>%
  filter(gas == "Helium",
         between(length, 20, 30),
         str_detect(phase, "5"),
         diameter < 0.3,
         thickness == 0.25) %>% 
  rename(CAS = query)
head(RIs_filtered)

RIs_filtered %>% 
  group_by(CAS) %>% 
  summarise(mean_RI = mean(RI))

best_RIs <-
  RIs_filtered %>%
  group_by(CAS) %>% 
  filter(RI == median(RI)) %>% 
  filter(row_number() == 1) %>% 
  select(CAS, RI, reference)
best_RIs

cir_query("o-Xylene", "cas")
```

#using chemlook
```{r}
library(chemlook)

name_querey <- cl_query('1071-83-6', from = 'cas')
```

>note for when you return, so the webchem is kinda working but needs the CAS numbers, so you need to go export those files again, but add the CAS numbers and then fill in the gaps of the missing CAS numbers before the RI search can work.
