---
title: "Thermal_Desorption_XML_to_CSV"
format: html
---

This was written with the help of Microsoft CoPilot, the code was comprehended and edited by me, Ruth Williamson, and I take full responsibility for the use and product.

insert libraries
--------------------------------------------
```{r}
library(xml2)
library(dplyr)
library(purrr)
library(tibble)
library(readr)
library(stringr)
library(janitor)
```


--------------------------------------------
Function to turn xml into tibble

```{r}
xml_compounds_to_tibble <- function(path,
                                   record_xpath = ".//Compounds",
                                   col_types = NULL,
                                   add_source = TRUE) {
  doc <- read_xml(path)

  nodes <- xml_find_all(doc, record_xpath)

  rows <- map(nodes, function(node) {
    kids <- xml_children(node)
    nms  <- xml_name(kids)
    vals <- xml_text(kids)

    # Normalize empty strings to NA
    vals <- na_if(str_trim(vals), "")

    as.list(setNames(vals, nms))
  })

  out <- bind_rows(rows) |> as_tibble()

  # Convert types if provided (keeps unspecified columns as character)
  if (!is.null(col_types)) {
    out <- type_convert(out, col_types = col_types, na = c("", "NA"))
  }

  if (add_source) {
    out <- mutate(out, source_file = basename(path), .before = 1)
  }

  out
}
```

--------------------------------------------
Extracts Meta-Data from Paths and Filenames

```{r}
extract_metadata <- function(path) {

  path_parts <- str_split(path, .Platform$file.sep)[[1]]

  file_dir    <- path_parts[length(path_parts) - 3]
  project_dir <- path_parts[length(path_parts) - 2]
  sample_dir  <- path_parts[length(path_parts) - 1]

  xml_name <- tools::file_path_sans_ext(basename(path))

  # ---- sample type ----
  sample_type <- case_when(
    str_detect(sample_dir, "^p\\d+")    ~ "sample", #Looks for samples if they have pump info
    str_detect(sample_dir, "^blk")      ~ "blk", #labels blanks by looking for something with "blk" in the name
    str_detect(sample_dir, "^control") ~ "control", #looks for controls if they are labeled "control"
    TRUE ~ NA_character_
  )

  # ---- shared metadata ----
  pump <- if_else(
    sample_type == "sample",
    str_extract(sample_dir, "p\\d+"),
    NA_character_
  )

   time = if_else(
      sample_type %in% c("sample", "control"),
      str_extract(sample_dir, "(?<=_)\\d{3,4}(?=_)|(?<=_)\\d{3,4}$"),
      NA_character_
      )
#edit these for your own specific variables you want to keep track of   
  location <- str_extract(project_dir, "como|forest|field")
  season   <- str_extract(project_dir, "spring|summer|fall|winter")
  year     <- str_extract(project_dir, "\\d{4}")
  mix      <- str_extract(xml_name, "mix\\d+")
  blk_num <- str_extract(sample_dir, "blk_\\d+")

  # ---- sample ID (type-aware) ----
  sample_ID <- case_when(

    sample_type == "sample" ~ str_c(
      "sample", pump, time, location, season, year, mix,
      sep = "_"
    ),

    sample_type == "control" ~ str_c(
      "control", time, location, season, year, mix,
      sep = "_"
    ),

    sample_type == "blk" ~ str_c(
      blk_num, location, season, year, mix,
      sep = "_"
    ),
    TRUE ~ NA_character_)

  tibble(
    project   = project_dir,
    type      = sample_type,
    pump      = pump,
    blk_num   = blk_num,
    time      = time,
    location  = location,
    season    = season,
    year      = year,
    mix       = mix,
    sample_ID = sample_ID
  )
}
```

--------------------------------------------
Reading all xml files

```{r}
read_all_xmls <- function(root_dir,
                          record_xpath = ".//Compounds",
                          col_types = NULL) {

  xml_files <- list.files(
    root_dir,
    pattern = "\\.xml$",
    recursive = TRUE,
    full.names = TRUE
  )

  message("Found ", length(xml_files), " XML files")

  map_dfr(xml_files, function(path) {

    compounds <- xml_compounds_to_tibble(
      path,
      record_xpath = record_xpath,
      col_types = col_types,
      add_source = TRUE
    )

    meta <- extract_metadata(path)

    bind_cols(
      meta |> slice(rep(1, nrow(compounds))),
      compounds
    )
  })
}
```

--------------------------------------------
column type assignment

```{r}
#can edit this to change column types (may want to change time to numeric?)

compound_types <- cols(
  Number                  = col_integer(),
  Peak_Name               = col_character(),
  Retention_Time          = col_double(),
  Amount                  = col_double(),
  Peak_Area               = col_double(),
  Units                   = col_character(),
  Cal_Istd                = col_integer(),
  QValue                  = col_integer(),
  Expected_Retention_Time = col_double(),
  Target_Ion              = col_double(),
  Qualifier1_Ion          = col_double(),
  Qualifier2_Ion          = col_double(),
  Qualifier3_Ion          = col_double(),
  Start_Integration       = col_double(),
  End_Integration         = col_double()
)


```

--------------------------------------------
put file here

```{r}
phytoncide_dat <- read_all_xmls( #put the name of your data table here
  root_dir = "phytoncide_collection_data_raw_xmls/", #put your data file you want to read in here
  col_types = compound_types) |> 
  clean_names() |> #cleans column names
  mutate(peak_name = peak_name |>
           str_trim () |>
           str_to_lower() |>
           str_replace_all(" ", "_")) #cleans the compound names

glimpse(phytoncide_dat)
```

------------------------------------------
Write ya CSV out!
```{r}
#write_csv(phytoncide_dat, "phytoncide_dat") #uncomment this line when you are ready to get your csv
```

-------------------------------------------
Sanity Checking that the data made it in correctly

```{r}
#check that rows match:
expected_rows <- (16*35)*3+(17*35)*3

if (nrow(foresttest) != expected_rows) {
  warning("Row count does not match expected value!")
}

#checking the compound counts per mix. Should be 10 for mixes 1-3 and 5 for mix 4.
compound_counts <- foresttest |>
  count(sample_ID, name = "n_compounds")

compound_counts

#can run unique on rows to make sure there aren't any typos or strange errors.
unique(foresttest$location)

unique(foresttest$type)

unique(foresttest$year)

unique(foresttest$mix)

```

It is also good practice to randomly check three or more lines of the tibble
with their original xml files to make sure things are matching up correctly

Once you are happy here, go ahead and uncomment the write_csv line and have fun with your data set!!!
